---
title: "Warpbreaks Report"
author: "STA303 In Class"
date: '2018-07-11'
output: pdf_document
---

```{r setup, include=FALSE,warning = FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
library(faraway)

# Load data
data(warpbreaks)
warpbreaks_tbl <- dplyr::as_data_frame(warpbreaks)

```

Welcome to the Warpbreaks data report! Anything we type here shows up as text.

We can add a title using a hashtag (or "pound sign" if you're over 35)

# View the Data

```{r view-data-1}
glimpse(warpbreaks_tbl)

# Check the observation counts in each level of the factors
warpbreaks_tbl %>%
  group_by(wool,tension) %>%
  summarize(cnt = n())
```

If you want to run code without showing the code, use `echo = FALSE`:

```{r view-data-2,echo = FALSE}
warpbreaks_boxplot_1 <- warpbreaks_tbl %>%
  ggplot(aes(x = wool,y = breaks)) +
  theme_classic() +
  geom_boxplot() +
  labs(title = "Boxplot of breaks by wool type",
       x = "Wool Type",
       y = "Breaks")

warpbreaks_boxplot_2 <- warpbreaks_tbl %>%
  ggplot(aes(x = tension,y = breaks)) +
  theme_classic() +
  geom_boxplot() +
  labs(title = "Boxplot of breaks by level of tension",
       x = "Tension Level",
       y = "Breaks")

cowplot::plot_grid(warpbreaks_boxplot_1,warpbreaks_boxplot_2,nrow=1)

```

Check out the interaction plot. We'll use a subtitle to identify this section:

## Interaction plots

Control the figure size with `fig.height` and `fig.width`:
```{r interaction-plot-1,fig.height=2,fig.width=2}

warpbreaks_interactionplot_1 <- warpbreaks_tbl %>%
  group_by(wool,tension) %>%
  summarize(group_mean = mean(breaks)) %>%
  ggplot(aes(x = wool,y = group_mean,group = tension)) +
  theme_classic() +
  facet_grid(~tension) +
  geom_point() +
  geom_line() +
  labs(title = "Interaction Plot, wool x tension",
       x = "Wool",
       y = "Mean # of breaks")

warpbreaks_interactionplot_2 <- warpbreaks_tbl %>%
  group_by(wool,tension) %>%
  summarize(group_mean = mean(breaks)) %>%
  ggplot(aes(x = tension,y = group_mean,group = wool)) +
  theme_classic() +
  facet_grid(~wool) +
  geom_point() +
  geom_line() +
  labs(title = "Interaction Plot, tension x wool",
       x = "Tension",
       y = "Mean # of breaks")

cowplot::plot_grid(warpbreaks_interactionplot_1,warpbreaks_interactionplot_2)

```

Objects you create in previous chunks are usable in future chunks:

```{r interaction-plot-2,fig.height=5,fig.width=5}
cowplot::plot_grid(warpbreaks_interactionplot_1,warpbreaks_interactionplot_2)
```

### Assess normality assumption

```{r normality-1}
warpbreaks_tbl %>%
  mutate_at("breaks",funs( (. - mean(.)) / sd(.))) %>%
  arrange(breaks) %>%
  mutate(q = qnorm(1:n() / (n() + 1))) %>%
  ggplot(aes(x = q,y = breaks)) +
  theme_classic() +
  geom_point() +
  geom_abline(slope = 1,intercept = 0) +
  labs(title = "Normal QQ-plot",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles")

```
